---
title: "nes-lter-nutrients-transect-emlassemblyline"
author: "Jaxine Wolfe"
date: "Oct 20, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE)

# clear workspace for local development
rm(list = ls())

# install libraries for EML assembly
# install.packages("devtools")
# install_github("EDIorg/EMLassemblyline")

# define R packages to require
libs <- c("tidyverse", "readxl", "lubridate", "devtools", "EMLassemblyline", "EML")
# load libraries
lapply(libs, require, character.only = TRUE)

# assign resused directory to a variable
dir <- "/Users/jaxinewolfe/Documents/WHOI/NESLTER/nes-lter-growth-grazing-chl/"
# set working directory
setwd(dir)
```

## Preparing Original Data for EDI

• if theres a "-" in the cast or a "/" in the bottle, insert a "|".
• the missing values in numeric fields should be numeric
```{r}
# set filename to variable
growgraze_xls <- "NES-LTER_Growth-Grazing_SMD_2019_11_01"

# Load growth and grazing data from excel sheet
growgraze <- read_excel(path = paste0(dir, "/", growgraze_xls, ".xlsx"), 
                  sheet = 1, na = "NA")

# DATA CLEANING
# standardize the formatting of the cast and bottle columns
growgraze$cast <- trimws(gsub('-', '|', growgraze$cast))
growgraze$bottle <- trimws(gsub('/', '|', growgraze$bottle))

# Add categorical columns to expand missing value explantions
# define columms containing missing values
mv_cols <- c("mu_N", "mu_N_Std", "grazing", "grazing_Std", "mu_10d", 
             "mu_10d_Std", "grazing_10d", "grazing_10d_Std", "mu_10u", 
             "mu_10u_Std", "grazing_10u", "grazing_10u_Std")

mv <- growgraze %>% select(mv_cols) %>%
  # rename to incidate that these are the missing value columns
  rename_at(mv_cols, ~ paste0(mv_cols, "_mv"))

# convert numerical values in missing value explanation columns to not missing 
mv[mv != "n/n" & mv != "n/a" & 
               mv != "n/d" & mv != "n/c"] <- "not missing"

# convert missing values in original columns to -9999
growgraze[growgraze == "n/n" | growgraze == "n/a" |
            growgraze == "n/d" | growgraze == "n/c"] <- -9999
# assign columns to numerical
growgraze[mv_cols] <- sapply(growgraze[mv_cols], as.numeric)
# str(growgraze) # check

# add the missing value explanation columns to the data
growgraze_EDI <- cbind(growgraze, mv)

# rearrange columns
growgraze_EDI <- growgraze_EDI[,c("project", "cruise", "date_time_UTC", "station", "cast", "bottle", "latitude", "longitude", "depth", "temp", "sal", "date_time_start_UTC", "date_time_end_UTC", "light_treatment", "mu_0", "mu_0_Std",
                                  "mu_N", "mu_N_mv", "mu_N_Std", "mu_N_Std_mv", "grazing", "grazing_mv", "grazing_Std", "grazing_Std_mv", "mu_10d",
                                  "mu_10d_mv",  "mu_10d_Std", "mu_10d_Std_mv" , "grazing_10d", "grazing_10d_mv", "grazing_10d_Std", "grazing_10d_Std_mv",
                                  "mu_10u", "mu_10u_mv", "mu_10u_Std", "mu_10u_Std_mv", "grazing_10u", "grazing_10u_mv", "grazing_10u_Std", 
                                  "grazing_10u_Std_mv", "temp_light_inc_filename")]

write.csv(growgraze_EDI, "nes-lter-growth-grazing-chl.csv", row.names = FALSE)
```

## Quality Assurance: Nutrients Check

This chunk checks the nutrient values used by Pierre against the nutrient values from teh NES-LTER nutrients transect. The second version that Pierre provided produced no mismatched averages.
```{r}

# load in nutrients data (provided by Pierre)
nutrients_xls <- "NES-LTER_Nutrients_SMD_2019_11_01.xlsx"
# Load in nutrient data from excel sheet
nut_pierre <- read_excel(path = paste0(dir, "/", nutrients_xls), 
                  sheet = 1, na = "NA")

# load in nes-lter nutrient transect data 
# these will be compared against the NES-LTER nutrient-transect data
nut_transect <- read_csv("https://raw.githubusercontent.com/WHOIGit/nes-lter-nutrient-transect/master/nes-lter-nutrient-transect.csv")

nut_transect_avg <- nut_transect %>%
  group_by(cruise, cast, niskin, depth) %>%
  # calculate the average nutrients across the replicate samples
  summarize(avg_nitrate_nitrite = mean(nitrate_nitrite),
            avg_phosphate = mean(phosphate),
            avg_silicate = mean(silicate),
            avg_ammonium = mean(ammonium)) %>%
  # isolate the surface samples for each cruise-cast combo
  group_by(cruise, cast) %>%
  slice(which.min(depth))

# merge datasets for comparison 
nut_check <- left_join(nut_pierre, nut_transect_avg,
                       by = c("cruise", "castNut"="cast", "depthNut"="depth"))
# rename nitrate nitrite col
colnames(nut_check)[colnames(nut_check)=="Nitrate+Nitrite"] <- "Nitrate_Nitrite"

# create empty columns to check mismatches
nut_check$nitrate_nitrite_match <- FALSE
nut_check$phosphate_match <- FALSE
nut_check$ammonium_match <- FALSE
nut_check$silicate_match <- FALSE

for (i in 1:nrow(nut_check)) {
    # if the average values match for each nutrient, set name to true
    if (round(nut_check$Nitrate_Nitrite[i], 2) == round(nut_check$avg_nitrate_nitrite[i], 2)) {
    nut_check$nitrate_nitrite_match[i] <- TRUE
    }
    if (round(nut_check$Phosphate[i], 2) == round(nut_check$avg_phosphate[i], 2)) {
    nut_check$phosphate_match[i] <- TRUE
    }
    if (round(nut_check$Ammonium[i], 2) == round(nut_check$avg_ammonium[i], 2)) {
    nut_check$ammonium_match[i] <- TRUE
    }
    if (round(nut_check$Silicate[i], 2) == round(nut_check$avg_silicate[i], 2)) {
    nut_check$silicate_match[i] <- TRUE
    }
  }

# isolate mismatches
mismatch <- nut_check[nut_check$silicate_match==FALSE,]

```

## Quality Assurance: Consistency Check
```{r}

gg_check_EN608 <- growgraze_EDI %>%
  select(project:sal) %>%
  filter(cruise == "EN608")
unique(gg_check_EN608)

gg_check_EN617 <- growgraze_EDI %>%
  select(project:sal) %>%
  filter(cruise == "EN617")
unique(gg_check_EN617)

gg_check_EN627 <- growgraze_EDI %>%
  select(project:sal) %>%  
  filter(cruise == "EN627")
unique(gg_check_EN627)

```


## Quality Assurance: Map Sampling Locations

```{r, echo=FALSE}
library(maps)
library(dplyr)

nes <- map_data("state") %>% filter(long > -72 & lat < 42)

# Map given coordinates
ggplot() +
  geom_polygon(data = nes, mapping = aes(x = long, y = lat, group = group),
               fill = NA, color = "grey50") +
  geom_point(growgraze, mapping = aes(x = longitude, y = latitude, color = cruise),
            size = 1) + 
  coord_fixed(1.3) +
  theme_classic()
```

## Create Template .txt files for EMLassembly

```{r}

# define title of EDI-ready file
growgraze_file <- "nes-lter-growth-grazing-chl"

# read in metadata info
headers <- read_excel(path = paste0(dir, "/", growgraze_xls, ".xlsx"), 
                  sheet = "ColumnHeaders", na = "NA")
personnel <- read_excel(path = paste0(dir, "/", growgraze_xls, ".xlsx"), 
                  sheet = "Personnel", na = "NA")
keywords <- read_excel(path = paste0(dir, "/", growgraze_xls, ".xlsx"), 
                  sheet = "Keywords", na = "NA")
custom_units <- read_excel(path = paste0(dir, "/", growgraze_xls, ".xlsx"), 
                  sheet = "CustomUnits", na = "NA")
catvars <- read_excel(path = paste0(dir, "/", growgraze_xls, ".xlsx"), 
                  sheet = "CategoricalVariables", na = "NA")

# write info data frames to text files for EML assembly
write.table(personnel,"personnel.txt", 
            quote = FALSE, na = "", sep = "\t", row.names = FALSE)
write.table(headers, paste0("attributes_", growgraze_file, ".txt"), 
            quote = FALSE, na = "", sep = "\t", row.names = FALSE)
write.table(keywords,"keywords.txt", 
            quote = FALSE, na = "", sep = "\t", row.names = FALSE)
write.table(custom_units,"custom_units.txt", 
            quote = FALSE, na = "", sep = "\t", row.names = FALSE)
write.table(catvars, paste0("catvars_", growgraze_file, ".txt"), 
            quote = FALSE, na = "", sep = "\t", row.names = FALSE)

```

## Construct XML for Nutrient Data

```{r}

growgraze_pkg <- "knb-lter-nes.5.1"

# Import EDI templates for diet dataset licensed under CCBY
import_templates(path = dir,
                 license = "CCBY",
                 data.files = paste0(growgraze_file, ".csv"))
# this will not overwrite existing files

# eliminate unneccesary templates
# bounding_boxes.txt must be eliminated to specify coverage manually
unlink(c("bounding_boxes.txt", "additional_info.txt"))

# Temporal coverage
# Will need this in make_eml YYYY-MM-DD
column_of_dates <- as.Date(growgraze_EDI$date_time_UTC)
startdate <- min(column_of_dates, na.rm = TRUE)
enddate <- max(column_of_dates, na.rm = TRUE)
# temporal.coverage argument is expecting objects of 'character' class, not 'Date'
startdate_as_character <- as.character(startdate)
enddate_as_character <- as.character(enddate)

# Geographic coverage
# Will need this order in make_eml: North, East, South, West
North <- max(growgraze_EDI$latitude, na.rm = TRUE)
East <- max(growgraze_EDI$longitude, na.rm = TRUE)
South <- min(growgraze_EDI$latitude, na.rm = TRUE)
West <- min(growgraze_EDI$longitude, na.rm = TRUE)

# ensure metadata template txt files are UTF-8 and at specified path
make_eml(path = dir,
         dataset.title = "Phytoplankton growth and microzooplankton grazing rates from NES-LTER transect cruises EN608, EN617, EN627.",
         data.table = paste0(growgraze_file, ".csv"),
         data.table.description = "Phytoplankton growth and microzooplankton grazing rates data cleaned for EDI",
         other.entity = c(paste0(growgraze_xls, ".xlsx"), "NES-LTER_temp_light_incubation_SMD.zip"),
         other.entity.description = c("Phytoplankton growth and microzooplankton grazing rates contributed by the Menden-Deuer lab",
                                      "Temperature and light intensity recorded during the incubation experiments"),
         temporal.coverage = c(startdate_as_character, enddate_as_character),
         geographic.description = "NES-LTER Transect",
         geographic.coordinates = c(North, East, South, West),
         maintenance.description = "ongoing",
         user.id = "NES",
         affiliation = "LTER",
         package.id = growgraze_pkg)

```

## Insert Custom Project Node

Step 1: Read in the xml file and the new xml node
Step 2: Insert the new node or replace the old node with the new node
Step 3: Write the modified xml file 

Instructions for node replacement using the xml2 package can be found [here](https://cran.r-project.org/web/packages/xml2/vignettes/modification.html).

```{r}
# install.packages("xml2")
library(xml2)

# Function inserts project node after the methods node of an xml document
project_insert <- function(x, projectnode) {
  if (is.na(xml_find_first(x, ".//project")) == FALSE) {
    stop("<project> node already exists")
  }
  
  # find methods node
  methodsnode <- xml_find_first(x, ".//methods")
  # add project node after methods and before dataTable
  xml_add_sibling(methodsnode, projectnode, where = "after")
  
  if (eml_validate(x) == FALSE) {
    warning("XML document not valid")
  }
  return(x)
}

# read in project node
project_eml <- read_xml("parent_project.txt", from = "xml")
# read in xml files exported by make_eml
growgraze_eml <- read_xml(paste0(dir, "/", growgraze_pkg, ".xml"), from = "xml")
# all objects should be of class c("xml_document" "xml_node")

# make the neccessary modifications
growgraze_eml_final <- project_insert(growgraze_eml, project_eml)

# write modified xml file 
write_xml(growgraze_eml_final, paste0(dir, "/", growgraze_pkg, ".xml"))

```

