---
title: "nes-lter-nutrients-transect-emlassemblyline"
author: "Kate Morkeski, Jaxine Wolfe"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE)

# install and load libraries
#library(devtools)
#install_github('WHOIGit/ediutilities')
#install_github("EDIorg/EMLassemblyline")

library(EMLassemblyline)
library(ediutilities)
library(here)
library(tidyverse)
library(readr)
library(EML)
library(xml2)

```

## Read in Data Table

```{r}

grazing_data <- read_csv("NES-LTER-chla-grazing-experiments-rates-qc-noquote.csv")

grazing_data$mu_N <- as.numeric(grazing_data$mu_N)
grazing_data$mu_N_std <- as.numeric(grazing_data$mu_N_std)

# write csv with missing values as NaN
write.csv(grazing_data, here("nes-lter-growth-grazing-chl.csv"), na = "NaN", row.names = FALSE)

```
## QA: Map Sampling Locations

Call the map_locs function from ediutilities to map the sampling locations. Perform a visual check.

```{r, echo=FALSE}

# Map Check
map_locs(df = grazing_data, xvar = "longitude", yvar = "latitude",
         region = "transect", colorvar = "cruise")

```
## Generate basic summary of data table

```{r}
# Just for inspecting the summary: change all character columns to factor
DF <- grazing_data
DF[sapply(DF, is.character)] <- lapply(DF[sapply(DF, is.character)], as.factor)
summary(DF)

unique(DF$cruise)
unique(DF$station)

```
## Compare to chlorophyll transect rosette data

```{r} 

chlor <- read_csv("NES-LTER-chla-T0-SMD.csv")

chl_avg <- chlor %>%
group_by(cruise, cast, niskin, filter_size) %>%
  summarise(chl = mean(chl)) %>%
  mutate(chl = round(chl, 2))

grazing_chl <- left_join(grazing_data, chl_avg, by = c("cruise", "cast", "niskin", "size_fraction" = "filter_size"))

grazing_chl <- grazing_chl %>%
  select(-date_time_utc_start, -date_time_utc_end, -temperature_incubation_avg, -temperature_incubation_std, -duration_incubation, -dilution, -mu_0, -mu_0_std, -mu_N, -mu_N_std, -grazing, -grazing_std, -iode_quality_flag)  %>% 
  mutate(chl_diff = Chla - chl) 

ggplot(grazing_chl, aes(x=as.factor(size_fraction), y=chl, color = "chlorophyll")) + 
  geom_boxplot() + xlab('size_fraction') + geom_boxplot(aes(x = as.factor(size_fraction), y = Chla, color = "grazing"))

ggplot(grazing_chl, aes(x=as.factor(size_fraction), y=Chla)) + 
  geom_boxplot() + xlab('size_fraction')
ggsave("grazing-chl-by-filter-size.png")

ggplot(grazing_chl, aes(x=as.factor(size_fraction), y=chl)) + 
  geom_boxplot() + xlab('size_fraction')
ggsave("chl-chl-by-filter-size.png")

summary(grazing_chl)

chl_diff_summary <- grazing_chl %>%
group_by(size_fraction) %>%
  summarise(chl_diff_mean = mean(chl_diff), 
            chl_diff_min = min(chl_diff),
            chl_diff_max = max(chl_diff)) 


```


## EML Assembly

This chunk outputs the final xml file for EDI through the following steps:

Step 1: Populating EML Assembly Line templates with metadata
Step 2: Calculating the geospatial and temporal coverage 
Step 3: Making the XML file 
Step 4: Inserting a custom NES-LTER parent project node 

```{r}

# define input files
edi_filename <- "nes-lter-growth-grazing-chl"
metadata <- "nes-lter-growth-grazing-chl-info"
pkg_id <- "knb-lter-nes.5.2"

# Make EML Templates 
# Specify methods file type as markdown
excel_to_template(here(metadata), edi_filename, rights='CCBY', file_type=".md")

# Data Coverage
# isolate date and geospatial columns for input
date_col <- as.Date(grazing_data$date_time_utc_sampling)
#date_col <- grazing_data$date_time_utc_sampling
lat_col <- grazing_data$latitude
lon_col <- grazing_data$longitude
# run function to determine geospatial and temporal coverage
coverage <- data_coverage(dates = date_col, lat = lat_col, lon = lon_col)

# Make EML
make_eml(here(),
         dataset.title = "Phytoplankton growth and microzooplankton grazing rates from NES-LTER transect cruises, ongoing since 2018.",
         data.table = paste0(edi_filename, ".csv"),
         data.table.description = "Phytoplankton growth and microzooplankton grazing rates with size-fractionated chlorophyll-a concentration",
         other.entity = c("chl-grazing-experiment-temp-inc-data.zip"),
         other.entity.description = c("Temperature and light intensity recorded during the incubation experiments"),
         temporal.coverage = c(coverage$startdate, coverage$enddate),
         geographic.description = "NES-LTER Transect",
         geographic.coordinates = c(coverage$North, coverage$East, coverage$South, coverage$West),
         maintenance.description = "ongoing",
         user.id = "NES",
         user.domain = "LTER",
         package.id = pkg_id)

project_insert(pkg_id, filename='parent_project_RAPID.txt')

```


