---
title: "nes-lter-nutrients-transect-emlassemblyline"
author: "Jaxine Wolfe"
date: "Oct 20, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE)

# clear workspace for local development
rm(list = ls())

# assign resused directory to a variable
dir <- "/Users/jaxinewolfe/Documents/WHOI/NESLTER/nes-lter-growth-grazing-chl/"
# set working directory
setwd(dir)

# define source for functions developed for the EDI packaging workflow
source("edi-utilities.R")

# install and load libraries
# install.packages("devtools")
# install_github("EDIorg/EMLassemblyline")

# define R packages to require
libs <- c("tidyverse", "readxl", "lubridate", "devtools", "EMLassemblyline", "EML", "maps", "xml2")
# load libraries
lapply(libs, require, character.only = TRUE)

```

## Preparing Provided Data for EDI

```{r}
# set filename to variable
pierre_data <- "NES-LTER_Growth-Grazing_SMD_2019_11_01"
edi_filename <- "nes-lter-growth-grazing-chl"

# Load growth and grazing data from excel sheet
growgraze <- read_excel(path = paste0(dir, "/", pierre_data, ".xlsx"), 
                  sheet = "DataTable", na = "NA")
# round necessary columns to standard precision
growgraze$latitude <- round(growgraze$latitude, 5)
growgraze$longitude <- round(growgraze$longitude, 5)
growgraze$depth <- round(growgraze$depth, 3)

# Data Cleaning ----------

# standardize the formatting of the cast and bottle columns
# if theres a "-" in the cast or a "/" in the bottle, insert a "|"
growgraze$cast <- trimws(gsub('-', '|', growgraze$cast))
growgraze$bottle <- trimws(gsub('/', '|', growgraze$bottle))

# Add categorical columns to expand missing value explantions
# define columms containing missing values
gg_cols <- c("mu_N", "mu_N_Std", "grazing", "grazing_Std", "mu_10d", 
             "mu_10d_Std", "grazing_10d", "grazing_10d_Std", "mu_10u", 
             "mu_10u_Std", "grazing_10u", "grazing_10u_Std")
# isolate and rename missing value columns
mv <- growgraze %>% select(gg_cols) %>%
  rename_at(gg_cols, ~ paste0(gg_cols, "_mv"))

# convert numerical values in missing value explanation columns to not missing 
mv[mv != "n/n" & mv != "n/a" & 
               mv != "n/d" & mv != "n/c"] <- "not missing"

# convert missing values in original columns to -9999
growgraze[growgraze == "n/n" | growgraze == "n/a" |
            growgraze == "n/d" | growgraze == "n/c"] <- -9999
# assign columns to numerical
growgraze[gg_cols] <- sapply(growgraze[gg_cols], as.numeric)
# str(growgraze) # check
# round growgraze values to two decimals
growgraze[, c("mu_0", "mu_0_Std", gg_cols)] <- round(growgraze[, c("mu_0", "mu_0_Std", gg_cols)], 2)

# add the missing value explanation columns to the data
growgraze_EDI <- cbind(growgraze, mv)

# rearrange columns to match the order in the attributes table
growgraze_EDI <- growgraze_EDI[,c("project", "cruise", "date_time_UTC", "station", "cast", "bottle", "latitude", "longitude", "depth", "temp", "sal", "date_time_start_UTC", "date_time_end_UTC", "light_treatment", "mu_0", "mu_0_Std",
                                  "mu_N", "mu_N_mv", "mu_N_Std", "mu_N_Std_mv", "grazing", "grazing_mv", "grazing_Std", "grazing_Std_mv", "mu_10d",
                                  "mu_10d_mv",  "mu_10d_Std", "mu_10d_Std_mv" , "grazing_10d", "grazing_10d_mv", "grazing_10d_Std", "grazing_10d_Std_mv",
                                  "mu_10u", "mu_10u_mv", "mu_10u_Std", "mu_10u_Std_mv", "grazing_10u", "grazing_10u_mv", "grazing_10u_Std", 
                                  "grazing_10u_Std_mv", "temp_light_inc_filename")]

# write csv for EDI
write.csv(growgraze_EDI, paste0(edi_filename, ".csv"), row.names = FALSE)
```

## QA: Retainig Data Integrity

Check that the numerical values and their precision were retained during the data cleaning.

```{r}
# isolate growgraze data provided by pierre
gg_pierre <- read_excel(path = paste0(dir, "/", pierre_data, ".xlsx"), 
                  sheet = "DataTable", na = c("n/n", "n/a", "n/d", "n/c")) %>% select(gg_cols)
# convert NAs to numerical missing values
gg_pierre[is.na(gg_pierre)] <- -9999

# isolate growgraze columns from the cleaned data
gg_edi <- growgraze_EDI %>% select(gg_cols)

# loop through paired columns to check dor discrepancies between the data tables
for (i in 1:length(gg_cols)) {
  val_check <- gg_pierre[,i] == gg_edi[,i]
  print(unique(val_check)) # these should all print TRUE
}

```


## QA: Nutrients Check

Check that the averaged nutrient values used by Pierre match those provided by the REST API for the NES-LTER nutrients transect. UPDATE: The second version that Pierre provided produced no mismatched averages. Code commented out.

```{r}

# # load in nutrients data (provided by Pierre)
# nutrients_xls <- "NES-LTER_Nutrients_SMD_2019_11_01.xlsx"
# # Load in nutrient data from excel sheet
# nut_pierre <- read_excel(path = paste0(dir, "/", nutrients_xls), 
#                   sheet = 1, na = "NA")
# 
# # load in nes-lter nutrient transect data 
# # these will be compared against the NES-LTER nutrient-transect data
# nut_transect <- read_csv("https://raw.githubusercontent.com/WHOIGit/nes-lter-nutrient-transect/master/nes-lter-nutrient-transect.csv")
# 
# nut_transect_avg <- nut_transect %>%
#   group_by(cruise, cast, niskin, depth) %>%
#   # calculate the average nutrients across the replicate samples
#   summarize(avg_nitrate_nitrite = mean(nitrate_nitrite),
#             avg_phosphate = mean(phosphate),
#             avg_silicate = mean(silicate),
#             avg_ammonium = mean(ammonium)) %>%
#   # isolate the surface samples for each cruise-cast combo
#   group_by(cruise, cast) %>%
#   slice(which.min(depth))
# 
# # merge datasets for comparison 
# nut_check <- left_join(nut_pierre, nut_transect_avg,
#                        by = c("cruise", "castNut"="cast", "depthNut"="depth"))
# # rename nitrate nitrite col
# colnames(nut_check)[colnames(nut_check)=="Nitrate+Nitrite"] <- "Nitrate_Nitrite"
# 
# # create empty columns to check mismatches
# nut_check$nitrate_nitrite_match <- FALSE
# nut_check$phosphate_match <- FALSE
# nut_check$ammonium_match <- FALSE
# nut_check$silicate_match <- FALSE
# 
# for (i in 1:nrow(nut_check)) {
#     # if the average values match for each nutrient, set name to true
#     if (round(nut_check$Nitrate_Nitrite[i], 2) == round(nut_check$avg_nitrate_nitrite[i], 2)) {
#     nut_check$nitrate_nitrite_match[i] <- TRUE
#     }
#     if (round(nut_check$Phosphate[i], 2) == round(nut_check$avg_phosphate[i], 2)) {
#     nut_check$phosphate_match[i] <- TRUE
#     }
#     if (round(nut_check$Ammonium[i], 2) == round(nut_check$avg_ammonium[i], 2)) {
#     nut_check$ammonium_match[i] <- TRUE
#     }
#     if (round(nut_check$Silicate[i], 2) == round(nut_check$avg_silicate[i], 2)) {
#     nut_check$silicate_match[i] <- TRUE
#     }
#   }
# 
# # isolate mismatches
# mismatch <- nut_check[nut_check$silicate_match==FALSE,]

```

## QA: Consistency Across Light Treatments

Determine if there are differences in metadata across the two light treatments for each of the cruises. Slight differences in time observed for cruise EN608. There was one row missing which was expected: There was no 15% light experiment at EN617 station L10.

```{r}

cruiselist <- c("EN608", "EN617", "EN627")

for (k in 1:length(cruiselist)) {
  gg_check <- growgraze_EDI %>%
    select(project:sal) %>%
    filter(cruise == cruiselist[k])
  unique(gg_check)
  print(gg_check)
}

```


## QA: Map Sampling Locations

Call the map_locs function from edi-utility.R to map the sampling locations. Perform a visual check.

```{r, echo=FALSE}

# Map Check
map_locs(growgraze_EDI, longitude, latitude, region = "transect")

```

## EML Assembly

This chunk outputs the final xml file for EDI through the following steps:

Step 1: Populating EML Assembly Line templates with metadata
Step 2: Calculating the geospatial and temporal coverage 
Step 3: Making the XML file 
Step 4: Inserting a custom NES-LTER parent project node 

```{r}

# define input files
metadata <- "NES-LTER_Growth-Grazing_SMD_2019_11_01"
edi_filename <- "nes-lter-growth-grazing-chl"
pkg_id <- "knb-lter-nes.5.1"

# Make EML Templates 
xlsx_to_template(metadata.path = metadata, 
                 edi.filename = edi_filename, 
                 rights = "CCBY")

# Data Coverage
# isolate date and geospatial columns for input
date_col <- as.Date(growgraze_EDI$date_time_UTC)
lat_col <- growgraze_EDI$latitude
lon_col <- growgraze_EDI$longitude
# run function to determine geospatial and temporal coverage
coverage <- data_coverage(dates = date_col, lat = lat_col, lon = lon_col)

# Make EML
make_eml(path = getwd(),
         dataset.title = "Phytoplankton growth and microzooplankton grazing rates from NES-LTER transect cruises EN608, EN617, EN627.",
         data.table = paste0(edi_filename, ".csv"),
         data.table.description = "Phytoplankton growth and microzooplankton grazing rates data cleaned for EDI",
         other.entity = c(paste0(metadata, ".xlsx"), "NES-LTER_temp_light_incubation_SMD.zip"),
         other.entity.description = c("Phytoplankton growth and microzooplankton grazing rates contributed by the Menden-Deuer lab.",
                                      "Temperature and light intensity recorded during the incubation experiments. Each .csv file includes dateTime in UTC, Temp_In (temperature in the incubation tank), Light_In (light in the incubation tank), Temp_Out (temperature out the incubation tank) and Light_Out (light out the incubation tank) with a time step of 5 min."),
         temporal.coverage = c(coverage$startdate, coverage$enddate),
         geographic.description = "NES-LTER Transect",
         geographic.coordinates = c(coverage$North, coverage$East, coverage$South, coverage$West),
         maintenance.description = "ongoing",
         user.id = "NES",
         user.domain = "LTER",
         package.id = pkg_id)

# Insert Custom Project Node
project_insert(edi_pkg = pkg_id)
```
